---
title: "Visualize diversity patterns of PopGen skin microbiome"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction

Here I will visually explore the diversity of PopGen skin microbiome. I will work with the pre-computed (see folder before) diversity measures. I want to visualize obvious patterns between the diversity measures and obvious phenotypes.

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 5,
               fig.width = 4 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)

# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(vegan)
library(reshape2)
library(ape)
library(RColorBrewer)
library(phyloseq)
```

## Import data
```{r}
# Input folders
ps <- "/home/lsilva/IKMB/projects/skin.mgwas/results/1.data.harmonization/3.microbiome.harmonized/phyloseq.rarefied.rds" %>% 
  readRDS()
micro.r <- ps %>% otu_table() %>% data.frame()
meta <- ps %>%
  sample_data() %>% 
  data.frame() %>% 
  mutate(Site = case_when(Site == "Antecubital fossa" & 
                            Cohort == "KORA FF4" ~ "Antecubital fossa (K)",
                          Site == "Antecubital fossa" & 
                            Cohort == "PopGen" ~ "Antecubital fossa (P)",
                          TRUE ~ Site))
```


# Get distances

```{r}
micro.dis <- micro.r %>%
  vegdist("bray")
```

# Beta-diversity

Let's have a look at the beta diversity

## Calculate pcoa
```{r}
pcoa <- micro.dis %>% 
  pcoa()
pcoa$values[1:5,]
biplot.pcoa(pcoa)
```

## Visualize distance of samples
```{r}

to.plot.beta <- pcoa$vectors[,1:2] %>%
  data.frame() %>% 
  rownames_to_column("NGS_ID") %>% 
  left_join(meta) %>%
  group_by(Participant_ID)


ggpubr::ggscatterhist(
  to.plot.beta %>% 
    mutate(Site = factor(Site, levels = c("Dorsal forearm",
                                          "Volar forearm",
                                          "Antecubital fossa (P)",
                                          "Antecubital fossa (K)",
                                          "Forehead",
                                          "Retroauricular fold"))),
  x = "Axis.1", y = "Axis.2",
  color = "Site",
  alpha = 0.4,
  margin.params = list(fill = "Site", color = "darkgrey", size = 0.2),
  margin.plot = "boxplot",
  ggtheme = theme_few(),
)

```


# Fancy plot

```{r}
var.explained <- pcoa$values %>% 
  pull(Relative_eig) %>% 
  `*` (100)
library(ggpubr)
g.pcoa <- to.plot.beta %>% 
  mutate(Site = Site %>%
           gsub("Dorsal", "D.",.) %>% 
           gsub("Antecubital", "A.",.) %>% 
           gsub("Volar", "V.",.) %>% 
           gsub("Retroauricular", "R.",.)) %>% 
  mutate(Site = factor(Site, levels = c("D. forearm",
                                        "V. forearm",
                                        "A. fossa (P)",
                                        "A. fossa (K)",
                                        "Forehead",
                                        "R. fold"))) %>% 
  ggplot(aes(x = Axis.1, y = Axis.2, group = factor(Site))) +
  geom_point(aes(color = Site, fill = Site)) +
  theme(legend.position = "bottom",
        text = element_text(size = 20),
        legend.text = element_text(size = 18),
        axis.text = element_text(size = 20),
        legend.margin=margin(-1,0,-1,0),
        legend.box.margin=margin(-1,0,-1,0)) +
  labs(x= paste0("Axis 1 (", var.explained[1] %>% round(2), "%)"),
       y =  paste0("Axis 2 (", var.explained[2] %>% round(2), "%)"))

g.pcoa <- ggExtra::ggMarginal(g.pcoa, type="boxplot", groupFill = TRUE,groupColour = T) 
g.pcoa
```

# Save

```{r}
"to.plot.beta.rds" %>%  
  paste0(d.out, "/", .) %>% 
  saveRDS(to.plot.beta, .)

"var.explained.rds" %>%  
  paste0(d.out, "/", .) %>% 
  saveRDS(var.explained, .)
```



# Session information

```{r}
sessionInfo()
```
