---
title: "Merge and filter microbiome data from popgen and kora"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction

Filter is done by abundance and prevalence based by Site and cohort data

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
#v.01
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 6,
               fig.width = 5 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)
# # Color blind friendly pallet
# cbp1 <- c("#E69F00", "#56B4E9", "#009E73",
#           "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")
# 
# scale_fill_discrete <- function(...) {
#   scale_fill_manual(..., values = cbp1)
# }
# scale_colour_discrete <- function(...) {
#   scale_colour_manual(..., values = cbp1)
# }

# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(phyloseq)
library(metagMisc) # filtering options
```

## Import data
```{r}
d.in <- d.out
ps.popgen <- paste0(d.in, "/","phyloseq.popgen.rds") %>% 
  readRDS()
ps.kora <-  paste0(d.in, "/","phyloseq.kora.rds") %>% 
  readRDS()
microenv <- "~/IKMB/projects/skin.mgwas/common.data/microenv.tsv" %>% 
  read.delim(stringsAsFactors = F)
```


# Functions

## Function to plot  prevalence abundance

```{r}
plot.phylum = function(ps, name){
  prevdf = apply(X = otu_table(ps),
                 MARGIN = ifelse(taxa_are_rows(ps), yes = 1, no = 2),
                 FUN = function(x){sum(x > 0)})
  prevdf1 = data.frame(Prevalence = prevdf,
                       TotalAbundance = taxa_sums(ps),
                       tax_table(ps)@.Data)
  p <- ggplot(prevdf1, aes((TotalAbundance/sum(otu_table(ps)@.Data))*100,
                           (Prevalence / nsamples(ps))*100,color=Phylum)) +
    # Include a guess for parameter
    geom_hline(yintercept = 10, alpha = 0.5, linetype = 2) +
    geom_vline(xintercept = 0.1, alpha = 0.5, linetype = 2) +
    geom_point(size = 2, alpha = 0.7) +
    scale_x_log10() +  xlab("Abundance (%)") + ylab("Prevalence (% samples)") +
    facet_wrap(~Phylum) + theme(legend.position="none") + labs(title = name)
  
  return(p)
}
```

## Function to select ASVs 
Selection is performed based on the intersection of prevalence and abundance cut-offs calculated on total rarefied ASV tables.

```{r}
select.asv <- function(cut.off.core, pc.cutoff, micro.st){
  
  micro.1 <- micro.st
  
  #Set cutoff core
  min.participants <- nrow(micro.1) * cut.off.core
  print(min.participants)
  # Get ASVs that pass the core cutoff
  core <- micro.1 > 0
  core <- core %>% colSums()
  core <- core[core > min.participants]
  core <- names(core)
  # Set cutoff abundance
  abun <- colSums(micro.1)
  total.seqs <- sum(abun)
  min.abun <- (pc.cutoff * total.seqs) / 100
  print(min.abun)
  # Get ASVs that pass abundance cutoff
  abun <- colSums(micro.1)
  abun <- abun[abun > min.abun]
  abun <- names(abun)
  # Combine the results
  asv.to.keep.1 <- base::intersect(core, abun) %>% unique()
  
  d <- data.frame(ASV = asv.to.keep.1, stringsAsFactors = F) 
  return(d)
}
```


# Combine data from the cohorts

## Set name of cohorts
```{r}
sample_data(ps.popgen)$Cohort = "PopGen"
sample_data(ps.kora)$Cohort = "KORA FF4"
```

## Merge 

## Change names of taxa to sequence 

Before merging, let's converge taxa (ASV1, ASV2...) names to sequences, since this is use by merge to sum abundances


```{r}
identical(tax_table(ps.popgen)@.Data %>% data.frame() %>% pull(seqs) %>% names(), taxa_names(ps.popgen))

taxa_names(ps.popgen) <- tax_table(ps.popgen)@.Data %>%
  data.frame() %>% 
  pull(seqs)

taxa_names(ps.popgen) %>% head()
#
taxa_names(ps.kora) <- tax_table(ps.kora)@.Data %>%
  data.frame() %>% 
  pull(seqs)


# It performs full join, which means that any OTU that is specific of a single cohort are eliminated Lets
ps <- merge_phyloseq(ps.kora, ps.popgen)

#Because ASV names are now possibly repeated, let's rename them, and use them as taxa names


nr <- ps@tax_table@.Data %>%
  data.frame %>%
  nrow()

tax_table(ps)@.Data[, "ASV"] <- paste0("ASV",
                                       str_pad(string = 1:nr,
                                               width = nr %>%
                                                 as.character() %>%
                                                 str_length(),
                                               side = "left",
                                               pad = "0"))
# Can we use them as taxa names
tax_table(ps)@.Data[, "ASV"] %>% duplicated() %>% sum() == 0

taxa_names(ps) <- tax_table(ps)@.Data[, "ASV"]
```

### merge_phyloseq does it a full join.
```{r}
k <- tax_table(ps.kora)@.Data %>% data.frame() %>% select(seqs) %>% mutate(kora = 1)
p <- tax_table(ps.popgen)@.Data %>% data.frame() %>% select(seqs) %>% mutate(popgen = 1)
identical(full_join(k, p, by = "seqs") %>% nrow(), ps@tax_table@.Data %>% data.frame %>% nrow())
rm(k,p)
```

# Add index of Site and Cohort

Indexes will be based on the combination of Site and Cohort. This will facilitate loops and subsisting the data later on.

```{r}
sample_data(ps)$index <- sample_data(ps) %>% 
  data.frame() %>%
  mutate(index = paste(Site, Cohort) %>% 
           as.factor() %>% 
           as.numeric() %>% 
           as.character()) %>%
  pull(index)
```

# Add microenvironment

```{r}
sample_data(ps)$microenv <- sample_data(ps) %>% 
  data.frame() %>%
  left_join(microenv, by = "Site") %>% 
  pull(Microenvironment)
```


# Filter based on abundance and prevalence by Site and Cohort

## Set cut-offs
```{r}
#Prevalence
cut.off.core <- 0.10 #proportion
# Abundance
pc.cutoff <- 0.1 #percentage
```

## Rarefy data

Need to normalize the data for filtering

```{r}
ps.rare <- ps %>% rarefy_even_depth(sample.size = 10000, replace = F, trimOTUs = T) 
```

## Filter

```{r, fig.width = 14, fig.height = 14}
ps.sites <- list()

indexes <- sample_data(ps) %>% pull(index) %>% unique() %>% sort()

for (i in indexes){
  # Get subset of ps and ps.rare (used to calculate the asvs to keep)
  ps.temp <- ps %>% subset_samples(index == i) %>%
    prune_taxa(taxa_sums(.) > 0, .)
  
  ps.rare.temp <- ps.rare %>% subset_samples(index == i) %>%
    prune_taxa(taxa_sums(.) > 0, .)
  # Plot 
  title <- ps.temp %>% 
    sample_data() %>%
    data.frame() %>% 
    select(Site, Cohort) %>%
    unique %>% 
    mutate(name = paste(Site, Cohort)) %>% 
    pull (name)
  plot.phylum(ps.temp, paste("Before:", title)) %>% print()
  
  # Select ASVs to keep based on rarefied
  asvs <- select.asv(cut.off.core, 
                     pc.cutoff, 
                     otu_table(ps.rare.temp)@.Data %>% data.frame)
  
  ps.temp <- ps.temp %>% subset_taxa(ASV %in% asvs$ASV)
  
  #visualize filtering
  plot.phylum(ps.temp, paste("After:", title)) %>% print()
  
  # add data to the list
  ps.sites[[i]] <- ps.temp
  rm(ps.temp, title, asvs, ps.rare.temp)
}
```

## Merge all sites

```{r}
first <- indexes[1]
indexes.no.first <- indexes[-1]

ps.filtered <- ps.sites[[first]]
for (i in indexes.no.first){
  ps.filtered <- merge_phyloseq(ps.filtered, ps.sites[[i]])
}
print("Bofore filtering:")
ps
print("After filtering:")
ps.filtered

```

## Remove samples on post filtering
Filter samples which remained with less than 5K sequences

```{r}
ps.filtered <- ps.filtered %>% 
  prune_samples(sample_sums(.) > 5000,.) %>% 
  prune_taxa(taxa_sums(.) > 0, .)


plot.phylum(ps.filtered, "Filtered") %>% print()
```

## Update ASV names

```{r}
#Because ASV names are now possibly repeated, let's rename them, and use them as taxa names


nr <- ps.filtered@tax_table@.Data %>%
  data.frame %>%
  nrow()

tax_table(ps.filtered)@.Data[, "ASV"] <- paste0("ASV",
                                                str_pad(string = 1:nr,
                                                        width = nr %>%
                                                          as.character() %>%
                                                          str_length(),
                                                        side = "left",
                                                        pad = "0"))
# Can we use them as taxa names
tax_table(ps.filtered)@.Data[, "ASV"] %>% duplicated() %>% sum() == 0

taxa_names(ps.filtered) <- tax_table(ps.filtered)@.Data[, "ASV"]

```

# Make list of participants

## Per site

```{r}
dir.participants <- paste0(d.out, "/participant.list")
dir.create(dir.participants , showWarnings = F)

for (i in indexes){
  df <- ps.filtered %>%
    subset_samples(index == i) %>%
    sample_data() %>%
    data.frame() %>% 
    select(Genotype_ID, Site, Cohort, index)
  name <- df %>%
    select(Site, Cohort, index) %>% 
    unique() %>% 
    mutate(name = paste(index, abbreviate(Site), abbreviate(Cohort), sep = ".") %>% 
             str_to_lower() %>% 
             str_replace(" ", ".")) %>% 
    pull(name)
  write.table(df$Genotype_ID %>% unique(), paste0(dir.participants, "/", name, ".txt"), quote = F,sep = "\t", row.names = F, col.names = F) 
  
}
```
# Make list of participants per cohort
```{r}
for (i in ps.filtered %>% 
     sample_data() %>%
     pull(Cohort) %>% 
     unique()){
  df <- ps.filtered %>%
    subset_samples(Cohort == i) %>%
    sample_data() %>%
    data.frame() %>% 
    select(Genotype_ID, Cohort)
  name <- df %>%
    select(Cohort) %>% 
    unique() %>% 
    mutate(name = Cohort %>% 
             str_to_lower() %>% 
             str_replace(" ", "")) %>% 
    pull(name)
  write.table(df$Genotype_ID %>% unique(), paste0(dir.participants, "/", name, ".txt"), quote = F,sep = "\t", row.names = F, col.names = F) 
  
}

```


# Save new data

```{r}
saveRDS(ps.filtered, paste0(d.out, "/", "phyloseq.rds"))
```



# Session information

```{r}
sessionInfo()
```
