---
title: "Filter and select PopGen microbiome"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction


# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
#v.01
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 6,
               fig.width = 5 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)
# Color blind friendly pallet
cbp1 <- c("#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = cbp1)
}
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = cbp1)
}

# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(phyloseq)
```

## Import data
```{r}
d.in <- "~/IKMB/projects/skin.mgwas/popgen/results/2.combine.metadata.microbiome"
source("~/IKMB/projects/skin.mgwas/scripts/functions/get.phyloseq.R") #load function to convert my files to phyloseq
micro <- paste0(d.in, "/", "micro.rds") %>% 
  readRDS()

meta <- paste0(d.in, "/", "meta.rds") %>% 
  readRDS()  

tax <- paste0(d.in, "/", "tax.rds") %>% 
  readRDS()  
```


# Make phyloseq object
```{r}
ps <- get.phyloseq(micro, tax, meta)
rm(micro, meta, tax)
```

# Remove samples with less than 10 K sequences
```{r}
ps
ps <- ps %>% 
  prune_samples(sample_sums(.) > 10000,.) 
ps
```

# Remove samples based on metadata

## Remove participants with no information about site or was taken from chronic legions

```{r}
ps
ps <- ps %>% 
  subset_samples(!is.na(Site)) %>% 
  subset_samples(Site != "Chronic lesion") 
ps
```

## Select confounders

```{r}
sd <- ps %>% 
  sample_data() %>% 
  data.frame() %>% 
  mutate(Age = t4_examination_age) %>% 
  mutate(Sex = t14110_gender %>% str_to_lower()) %>% 
  mutate(BMI = t277_BMI) %>% 
  mutate(Site = Site %>% as.character()) %>%
  mutate(Genotype_ID = Participant_ID) %>% 
  select(Genotype_ID, Participant_ID, NGS_ID, Site, Age, Sex, BMI)
rownames(sd) <- sd$NGS_ID

sample_data(ps) <- sd %>% sample_data()
ps
rm(sd)
```

## Keep only complete data

```{r}
ps
ps <- ps %>% 
  subset_samples(., ps %>% 
                   sample_data() %>% 
                   complete.cases())
ps  
```


# Remove empty otus

```{r}
ps
ps <- ps %>% 
  prune_taxa(taxa_sums(.) > 0, .)
ps
```



# Save

```{r}
paste0(d.out, "/", "phyloseq.popgen.rds") %>% 
  saveRDS(ps, .)
```

# Session information

```{r}
sessionInfo()
```
