---
title: "Choice of the glm algorithm"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction
Should I use MASS or MVabund algorithm?

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
#v.01
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 6,
               fig.width = 5 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)
# Color blind friendly pallet
cbp1 <- c("#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = cbp1)
}
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = cbp1)
}

# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}

library(MASS)
library(tidyverse)
library(phyloseq)
library(mvabund)

```

## Function
```{r}
do.MASS <- function (testdata) {
    glm.nb(tax ~ Age + Sex + BMI + PCA1 + PCA2 + PCA3 + PCA4 + PCA5 + PCA6 + PCA7 + PCA8 + PCA9 + PCA10 + offset(log(library.size)), data=testdata) %>% 
      AIC() %>% 
      data.frame(formula = "MASS", AIC =.) %>%
      mutate(test = id.test) %>% 
      return()
  }
  
  do.mvabund <- function (testdata) {
    manyglm(tax ~ Age + Sex + BMI + PCA1 + PCA2 + PCA3 + PCA4 + PCA5 + PCA6 + PCA7 + PCA8 + PCA9 + PCA10 + offset(log(library.size)), data=testdata, family="negative.binomial") %>% 
      AIC() %>% 
      data.frame(formula = "mvabund", AIC =.) %>%
      mutate(test = id.test) %>% 
      return()
  }
```


#  Perform the loop


```{r}
res <-NULL 
for (i in 1:528){
  id.test <- i
  params <- readRDS("~/IKMB/projects/skin.mgwas/results/2.statistical.tests/2.taxonomic.features/tests.rds") %>% 
    filter(test.id == get("id.test"))
  
  
  level = params %>% pull(level) %>% as.character()
  tax = params %>% pull(tax) %>% as.character()
  index = params %>% pull(index) %>% as.character()
  ncore = params %>% pull(ncore) %>% as.numeric()
  
  # Import data ----
  
  # Import microbiome data
  
  ps <- paste0("./microbiome.data/index", index, ".ps.rds") %>% 
    readRDS()
  
  # Select metadata
  testdata <- ps %>% sample_data() %>% data.frame() %>% select(Age, Sex, BMI, PCA1, PCA2, PCA3, PCA4, PCA5, PCA6, PCA7, PCA8, PCA9, PCA10, Genotype_ID)
  
  # Add tax data
  e <- paste0(level, "==", "'",tax, "'")
  
  tax.data <- ps %>% tax_glom(level) %>% subset_taxa(eval(parse(text = e)))
  testdata$tax <- otu_table(tax.data)@.Data
  tax.name <- tax_table(tax.data)@.Data %>% data.frame(stringsAsFactors = F) %>% select_if(~ !any(is.na(.))) %>% .[1,] %>% paste(collapse = ";")
  # Add check
  identical(rownames(testdata), rownames(otu_table(tax.data)@.Data))
  
  
  # add offsets
  testdata$library.size = sample_sums(ps)
  
  # Get zero-truncated abundance of bacterial features and filter them for extreme outliers, deviating more than 5× the interquartile range (IQR) from the median abundance
  
  
  testdata <- testdata[testdata$tax > 0,]
  med <- median(testdata$tax)
  qt <- quantile(testdata$tax, c(0.25,0.75))
  iqr <- qt[2] - qt[1]
  testdata <- testdata[testdata$tax < (med + 5*iqr),]
  
  # Keep only complete cases
  
  testdata <- testdata[complete.cases(testdata),]
  
  print(paste("index", index, "level", level, "tax", tax, "nrow", nrow(testdata)))
  
  # Fit GLMs
  
  df <- tryCatch(do.MASS(testdata), error = function(e){data.frame(formula = "MASS", AIC = NA, test = id.test)})
  res <- rbind(res, df)
  
  df <- tryCatch(do.mvabund(testdata), error = function(e){data.frame(formula = "mvabund", AIC = NA, test = id.test)})
  res <- rbind(res, df)
}

```

# Visualize the results

```{r}
res2 <- res  %>% spread(formula, AIC) %>% mutate(diff = MASS - mvabund) %>% mutate(res = if_else(diff < 0,
                                                                                                 "MASS",
                                                                                                 "mvabund"))
res2 %>% count(res)

# 
ggpubr::ggscatter(res2 %>% filter(!is.na(MASS)), x = "MASS", y = "mvabund",
                  add = "reg.line",                                 # Add regression line
                  conf.int = TRUE,                                  # Add confidence interval
                  add.params = list(color = "blue",
                                    fill = "lightgray")) +
  ggpubr::stat_cor(method = "pearson", label.x = 3, label.y = 30) +
  ggpubr::stat_regline_equation()

res %>% filter(!is.na(AIC)) %>% count(formula)

```


# Conclusion

Mvabund has insignificantly slightly worst fitting. However, it is capable of fitting all the models, while the the MASS glm.nd does not. So I am fitting with mvabund.


# Session information

```{r}
sessionInfo()
```
