---
title: "Prepare data for tests in the cluster"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction

I will prepare data for paralellel tests inthe cluster

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 5,
               fig.width = 4 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)

# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(tidyverse)
library(phyloseq)
library(doParallel)
library(apcluster)
```

#Load 

```{r}
d.in <- "~/IKMB/projects/skin.mgwas/results/2.statistical.tests/2.taxonomic.features/microbiome.data/"
ps.r <- readRDS("~/IKMB/projects/skin.mgwas/results/1.data.harmonization/3.microbiome.harmonized/phyloseq.rarefied.rds")
```


# Parameters to test

```{r}
index = 1:6
ncore = 8
levels = c("Phylum" , "Class"  , "Order" ,  "Family" , "Genus" ,  "Species", "ASV")
```
# Select features that are present in both sites of each microenvironment

## Extract features from each site-cohort
```{r}
tests <- NULL
for (i in index){
  ps <- paste0(d.in,"index", i, ".ps.rds" ) %>% 
    readRDS()
  meta <- ps %>% 
    sample_data() %>% 
    data.frame() %>%
    select(Site, Cohort, index, microenv) %>%
    unique()
  tax.table <- tax_table(ps)@.Data %>% data.frame()
  for (z in levels){
    in.df <- data.frame(level = z,
                        tax = tax.table %>%
                          pull(z) %>%
                          as.character() %>% 
                          unique(),
                        meta)
    tests <- bind_rows(tests, in.df)
    rm(in.df)
  }
  
}
```

## Select

```{r}
nrow(tests)
tests <- tests %>% 
  group_by(level, tax, microenv) %>% 
  mutate(n = n()) %>% 
  filter(n == 2) %>% 
  ungroup()
nrow(tests)
```


# Select features based on their presence/abundance and median abundance

## Get stats

```{r}
tests <- data.frame(tests, median = NA, presence = NA)

ps.list <- list()
for (i in index){
  ps.list[[i]]<- paste0(d.in,"index", i, ".ps.rds" ) %>% 
    readRDS()
  names(ps.list)[i] <- i
}

# Do

for (i in 1:nrow(tests)){
  index.i <- tests[i,] %>% pull(index) %>% as.character()
  ps.i <- ps.list[[index.i]]
  level.i = tests[i,] %>% pull(level) %>% as.character()
  tax.i = tests[i,] %>% pull(tax) %>% as.character()
  
  # Get taxa info
  e <- paste0(level.i, "==", "'",tax.i, "'")
  tax.data <- ps.i %>% tax_glom(level.i) %>% subset_taxa(eval(parse(text = e)))
  feature.tax <- otu_table(tax.data)@.Data
  # Get taxa information
  m <- median(feature.tax[feature.tax>0])
  p <- sum(feature.tax > 0) 
  
  # save results
  tests[i,"median"] <- m
  tests[i,"presence"] <- p
}
```

## Visualize results

```{r}
ggplot(tests, aes(presence, median)) +
  geom_point(aes(color = level)) +
  facet_wrap(~ Cohort + Site) +
  scale_y_log10() +
  geom_hline(yintercept = 50, linetype = "dashed") +
  geom_vline(xintercept = 100, linetype = "dashed")
```

## Filter

```{r}
tests <- tests %>% 
  mutate(n = NA) %>% 
  filter(median > 50, presence > 100) %>% 
  group_by(level, tax, microenv) %>% 
  mutate(n = n()) %>% 
  filter(n == 2) %>% 
  ungroup()
```



## Stats: Summarize features
```{r}
s <- tests %>% select(level,tax) %>% unique() %>% group_by(level) %>% count(level)
#tota
sum(s$n)
s

tests %>% group_by(level, tax) %>% summarize(n = n()) %>% 
  count(n) %>% 
  ggplot(aes(x = n, y = nn)) +
  geom_col(aes(fill = level)) +
  facet_wrap(~ level) +
  theme(legend.position = "none")+
  labs(x = "Sites", y = "Counts")
```
# Cluster microenvironment-wide samples considering their similar non-truncated distribution

```{r}
cluster.features <- function (env, similarity.cutoff){
  #Isolate features to correlate
  tests.env <- tests %>% filter(microenv == env) %>% 
    select(level, tax) %>% 
    unique() %>% 
    mutate(name = paste(level, tax, sep = "_") %>% make.names(unique = T))
  
  # Filter rarefied phyloseq file
  samdata <- sample_data(ps.r) %>% data.frame()
  to.filter <- samdata$microenv == env
  ps.r.e <- ps.r %>% prune_samples(to.filter, .)
  identical(sample_data(ps.r.e) %>% data.frame() %>% pull(microenv) %>% unique, env) %>% 
    print()
  nsamples(ps.r.e) %>% print()
  
  # Get abundances of features
  print(paste("gathering data for", env))
  
  df <- data.frame(dummy = rep(1, nsamples(ps.r.e)))
  
  for (i in 1:nrow(tests.env)){
    # get variables
    level.i = tests.env[i,] %>% pull(level) %>% as.character()
    tax.i = tests.env[i,] %>% pull(tax) %>% as.character()
    name.i = tests.env[i,] %>% pull(name) %>% as.character()
        
    # Get taxa info - Have to do a workournd subset_taxa because it is bugging withing the function
    ps.r.e.i <- ps.r.e %>% tax_glom(level.i) 
    otu.i <- otu_table(ps.r.e.i)@.Data %>% data.frame()
    taxdata.i <- tax_table(ps.r.e.i)@.Data %>% data.frame()
    to.filter <- rownames(taxdata.i)[taxdata.i[,level.i] == tax.i] 
    feature.tax <- otu.i[, to.filter] %>% data.frame()
    colnames(feature.tax) <- name.i
    #gather results
    df <- cbind(df, feature.tax)
  }
  
  # Cluster using average
  df <- df %>% data.frame() %>% select(-dummy)
  
  dm <- 1- corSimMat(t(df), sel=NA, r=1, signed=F, method="spearman")
  
  hc <- hclust(as.dist(dm), method = "average")
  
  plot(hc, main = env)
  abline(h = 1 - similarity.cutoff, lty=2)
  
  hc.cut <- cutree(hc, h = 1 - similarity.cutoff) %>% 
    data.frame(cluster = .) %>% 
    rownames_to_column("name") %>% 
    mutate(cluster = paste(env, str_pad(cluster,
                                        width = 3, 
                                        side = "left", 
                                        pad = "0")))
  # Add annotation
  hc.cut <- hc.cut %>% left_join(tests.env, by = "name")
  return(hc.cut)
}
```

## Retrieve clustering information
```{r}
cutoff = 0.985
moist.c <- cluster.features(env = "Moist", similarity.cutoff = cutoff) %>% 
  mutate(microenv = "Moist")
seb.c <-  cluster.features(env = "Sebaceous", similarity.cutoff = cutoff) %>% 
  mutate(microenv = "Sebaceous")
dry.c <-  cluster.features(env = "Dry", similarity.cutoff = cutoff) %>% 
  mutate(microenv = "Dry")
```
## Join data

```{r}
clusters <- rbind(moist.c, seb.c, dry.c)
# modify name to human readable
clusters <- clusters %>% 
  mutate(name = paste0(str_sub(level,1,1) %>% str_to_lower(), ".", tax))
```

## Add order of levels

```{r}
order <- data.frame(level = levels, order = 1:7)
clusters <- clusters %>% left_join(order, by = "level")
```

## Get representative

```{r}
clusters.rep <- clusters %>% group_by(cluster) %>% 
  filter(order == max(order))
```

## Get All members of the cluster

```{r}
clusters.rep$cluster.member <- NA
for (i in 1:nrow(clusters.rep)){
  #Get cluster
  cluster.i <- clusters.rep[i,] %>% pull(cluster) %>% as.character()
  # Get all members of the cluster
  members <- clusters %>% filter(cluster == cluster.i) %>% 
    arrange(order) %>% 
    pull(name) %>% 
    paste(collapse = ";")
  clusters.rep$cluster.member[i] = members
  rm(members, cluster.i)
}
```

# Add information to tests

```{r}
tests.final <- inner_join(tests, clusters.rep, by = c("level", "tax", "microenv"))
```

## Stats: Summarize features
```{r}
s <- tests.final %>% select(level,tax) %>% unique() %>% group_by(level) %>% count(level)
#tota
sum(s$n)
s

tests %>% group_by(level, tax) %>% summarize(n = n()) %>% 
  count(n) %>% 
  ggplot(aes(x = n, y = nn)) +
  geom_col(aes(fill = level)) +
  facet_wrap(~ level) +
  theme(legend.position = "none")+
  labs(x = "Sites", y = "Counts")
```

# Add parameters

```{r}
tests.final <- tests.final %>% 
  ungroup %>% 
  arrange(cluster, Site) %>% 
  mutate(test.id = 1:n(),
         ncore := ncore)
```


# Save all dataset

```{r}
"tests.rds" %>% 
  paste0(d.out, "/", .) %>% 
  saveRDS(tests.final, .)
```

# Session information

```{r}
sessionInfo()
```
