---
title: "Gather results from single tests and finemapping"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      toc_float: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---
# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
#v.02
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 6,
               fig.width = 5 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)

# Set ggplot2 them as black and with
library("ggplot2")
theme_set(theme_bw())
# Set seed for reproducibility
set.seed(13)

# # Color blind friendly pallet
# cbp1 <- c("#E69F00", "#56B4E9", "#009E73",
#           "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")
# 
# scale_fill_discrete <- function(...) {
#    scale_fill_manual(..., values = cbp1)
# }
# scale_colour_discrete <- function(...) {
#   scale_colour_manual(..., values = cbp1)
# }

# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(tidyverse)
library(data.table)
```

## Import data
```{r}
fm.snps <- read_tsv("/work_ifs/sukmb447/projects/skin.mgwas/results/4.fine.mapping/1.meta/1.microenv/2.annotate.data/combined/all.snps.from.mapping.tsv")

# Add results directory
beta.dir <- "/work_ifs/sukmb447/projects/skin.mgwas/results/3.meta.analysis/1.microenv/1.beta.diversity/2.gather.and.filter"
feature.dir <- "/work_ifs/sukmb447/projects/skin.mgwas/results/3.meta.analysis/1.microenv/2.taxonomic.features/2.gather.and.filter"

# Add datasets indexes
beta.indexes <- data.frame(index.1 = c(1,3,4),
                           index.2 = c(2,6,5),
                           microenv = c("Moist", "Dry", "Sebaceous")) %>% 
  mutate(meta.id = paste0(index.1, index.2))
feature.indexes <- "/work_ifs/sukmb447/projects/skin.mgwas/results/3.meta.analysis/1.microenv/2.taxonomic.features/metasoft/1.metaanalysis/prepare/meta.indices.rds" %>% 
  readRDS()
```

# Get indexes to consider
Because tests and indexed, one for beta-diversity another for taxonomic features, I want to sort the indexes of the snps that resulted from finemapping, so I read only the files that I am interested in.

```{r}
fm.indexes <- fm.snps %>% 
  select(meta.id, metanalysis, type) %>% 
  distinct()
```
# Get indexes from tests

```{r}
beta.indexes <- beta.indexes %>% 
  mutate(index.1 = index.1 %>% as.character(),
         index.2 = index.2 %>% as.character(),
         meta.id = meta.id %>% as.numeric(),
         type = "beta",
         metanalysis = if_else(microenv == "Dry",
                               "comb",
                               "meta"))


feature.indexes <- feature.indexes %>% 
  mutate(type = "feature",
         metanalysis = if_else(microenv == "Dry",
                               "comb",
                               "meta"))
test.indexes <- bind_rows(beta.indexes, feature.indexes)
```

# Filter test.indexes
Keep only the indexes for the files that contain info about the finemapping snps.

```{r}
test.indexes <- left_join(fm.indexes, test.indexes,
                          by = c("meta.id", "metanalysis", "type"))
```

# Collect data for tests on taxonomic features sets
```{r}
# Function
get.data <- function(row.number){
  # Set paramenters
  i <- row.number #indexes row number
  type.i <- indexes[i,"type", drop = T]
  meta.id.i <- indexes[i,"meta.id", drop = T]
  test.1.i <- indexes[i, "test.1", drop = T]
  test.2.i <- indexes[i, "test.2", drop = T]
  index.1.i <- indexes[i, "index.1", drop = T]
  index.2.i <- indexes[i, "index.2", drop = T]
  
  # Read tests on features and betadiversity tests. They have different name structures and folders.
  
  if (type.i == "beta"){
    name.file <-  paste0(beta.dir, "/", "gathered.meta", meta.id.i, ".tsv.gz")
    
    df.i <- fread(name.file)
  }
  
  if (type.i == "feature"){
    name.file <- paste0("meta",
                        meta.id.i,
                        ".test", 
                        test.1.i,
                        ".test", 
                        test.2.i, 
                        ".index", 
                        index.1.i, 
                        ".index", 
                        index.2.i) %>%  
      paste0(feature.dir, "/", "parsed", ., ".txt.gz")
    
    df.i <- fread(name.file)
  }
  
  # Select snps of interest
  
  fm.snps.i <- fm.snps %>% 
    filter(meta.id == meta.id.i,
           type == type.i) %>% 
    select(snp.name)
  
  df.i <- left_join(fm.snps.i, df.i, by = "snp.name")
  
  return(df.i)
}


# Run beta
indexes <- test.indexes %>% filter(type == "beta")
beta.tests.results <- map(1:nrow(indexes), ~ get.data(.x)) %>% 
  bind_rows()

# Run features

indexes <- test.indexes %>% filter(type == "feature")
feature.tests.results <- map(1:nrow(indexes), ~ get.data(.x)) %>% 
  bind_rows()
```

# Save outputs

```{r}
# Beta
paste0(d.out, "/","gather.selected.beta.rds") %>% 
  saveRDS(beta.tests.results, .)

# Feature
paste0(d.out, "/","gather.selected.feature.rds") %>% 
  saveRDS(feature.tests.results, .)

```

# Session information

```{r}
sessionInfo()
```
