---
title: "Compare groups using DESeq2"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction

I will compare groups using deseq2. The script is divided into two parts. First, I have to format the input data and second I will perform deseq2 routine for all comparisons.

Raw counts were used as input for linear models with negative binomial distribution. DESeq2 analysis was performed with standard configurations, which includes a Wald test and p-value adjustedment for multiple testing. Effect sizes were shrank using the R package apeglm. Results were filtered for significance (adjusted p-values < 0.05) and effect size (log2fold change > 1). 

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
#v.01
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 6,
               fig.width = 5 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)
# Color blind friendly pallet
cbp1 <- c("#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = cbp1)
}
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = cbp1)
}

# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(biomaRt)
library(tidyverse)
library(SummarizedExperiment)
library(MatrixGenerics)
library(matrixStats)
library(DESeq2)
```

## Import data
```{r}
metadata <- read.csv("~/IKMB/projects/skin.mgwas/results/10.2d.rnaseq/0.prepare.data/rna.sample.sheet.csv")
rna <- readRDS("~/IKMB/projects/skin.mgwas/results/10.2d.rnaseq/1.nfrnaseq/star_salmon/salmon.merged.gene_counts.rds")
tests <- read_tsv("/home/lsilva/IKMB/projects/skin.mgwas/scripts/10.2d.rnaseq/2.deseq2/tests.tsv")
```


# Format input data

## Add metadata to rna experiment

### Get experiment data
```{r}
experiment.meta <- colData(rna) %>% 
  as.data.frame()
```

### Get metadata
```{r}
metadata <- metadata %>% 
  mutate(names = paste0(group, "_", "R", replicate))
```

### Combine them

```{r}
merged.meta <- experiment.meta %>% 
  left_join(metadata, by = "names")

rownames(merged.meta) <- merged.meta$files
identical(experiment.meta$files, merged.meta$files)
```

### Add to S4 object

```{r}
colData(rna) <- DataFrame(merged.meta)
```

## Add tests SummaziredExperiment

### Format
```{r}
tests <- tests %>% 
  mutate(index  = 1:nrow(.))
```
### Add to S4 object

```{r}
metadata(rna)$formula <- ~ group
metadata(rna)$tests <- tests
```

## Save 

```{r}
paste0(d.out, "/", "summarized.experiment.rds") %>% 
  saveRDS(rna, .)

```
# Annotate genes
```{r}
gene_mart <- useEnsembl(biomart = "ensembl", 
                        dataset = "hsapiens_gene_ensembl")
genes <- data.frame(Gene = rna@NAMES)

genes.annotation <- getBM(attributes = c("ensembl_gene_id",
                                           "external_gene_name",
                                           "description",
                                           "band",
                                           "gene_biotype"), 
                            filters = 'ensembl_gene_id', 
                            values = genes$Gene, 
                            mart = gene_mart) %>% 
    separate(description, c("description"), sep = " \\[" )

genes <- left_join(genes,
                   genes.annotation,
                   by = c("Gene" = "ensembl_gene_id")) %>% 
  dplyr::rename(Gene_name = external_gene_name) %>% 
  mutate(Gene_name = if_else(Gene_name == "" | is.na(Gene_name),
                             Gene, 
                             Gene_name))
```


# Perform analysis for each comparison

## Comparisons

```{r}
metadata(rna)$tests
```

## Perform deseq2 routine
```{r}
for (index.i in metadata(rna)$tests$index){
  
  # Extract parameters 
  params <- metadata(rna)$tests %>% filter(index == index.i)
  to.test <- c(params$group1,params$group2) 
  se <- rna[, rna$group %in% to.test]
  
  
  # Extract data from SummarizedExperiment (would not be necessary, but I am finding a bug in DESeqDataSet)
  
  countdata <- round(assays(se)[["counts"]])
  coldata <- colData(se) %>% 
    data.frame() %>% 
    mutate(group = factor(group, levels = to.test))
  
  # Make deseq object
  dds <- DESeqDataSetFromMatrix(countData = countdata,
                                colData = coldata,
                                design = metadata(se)$formula)
  # Perform test
  dds <- DESeq(dds)
  resultsNames(dds)[2]
  res <- results(dds, name = resultsNames(dds)[2])
  resLFC <- lfcShrink(dds, coef=resultsNames(dds)[2], type="apeglm")
  resOrdered <- resLFC[order(resLFC$pvalue),]
  
  #  Explore
  plotMA(resLFC, ylim=c(-2,2), main = resultsNames(dds)[2])
  abline(h= 1, lty = 2, col = "grey")
  abline(h= -1, lty = 2, col = "grey")
  
  ## Save results
  
  selected <- resLFC %>% 
    data.frame() %>%
    filter(abs(log2FoldChange) > 1, padj < 0.05) %>% 
    arrange(padj) %>% 
    bind_cols(params) %>% 
    mutate(test = resultsNames(dds)[2]) %>% 
    rownames_to_column("Gene") %>% 
    left_join(genes, by = "Gene") %>% 
    select(Gene, Gene_name, everything())
  
  all <- resLFC  %>% 
    data.frame() %>% 
    arrange(padj) %>% 
    bind_cols(params) %>% 
    mutate(test = resultsNames(dds)[2])  %>% 
    rownames_to_column("Gene") %>% 
    left_join(genes, by = "Gene") %>% 
    select(Gene, Gene_name, everything())
  
  paste0(d.out, "/", index.i,".", resultsNames(dds)[2],".filtered.tsv" ) %>% 
    write_tsv(selected, .)    
  paste0(d.out, "/", index.i,".", resultsNames(dds)[2],".all.tsv" ) %>% 
    write_tsv(all, .)  
}
```


# Session information

```{r}
sessionInfo()
```
